---
title: "ufo example"
author: "Haiyi Shi"
date: ""
output: md_document
---

## Modelling UFO locations as a log-Gaussian Cox process

#### Packages

```{r}
library(INLA)
library(maps)
library(sf)
library(sp)
library(ggplot2)
library(patchwork)
source("functions.r") ## source required functions
```


#### The Data and US map

```{r}
## read in data
ufo <- readr::read_csv("data/ufo.csv")
## us map
states <- map_data("state")
usa <- map_data('usa')
usa_region <- data.frame(Longitude = usa$long, Latitude = usa$lat)
```


Then, convert the existing lat-long coordinates to UTM (easting-northing system), using North America Lambert Conformal Conic.

```{r}
## define region and points
region <- as(sf::st_as_sf(maps::map("usa", fill = TRUE, plot = FALSE)), "Spatial")
locs <- data.frame(x = ufo$city_longitude, y = ufo$city_latitude)
ufo_locs <- ufo
coordinates(ufo_locs) <- c("city_longitude", "city_latitude")
proj4string(ufo_locs) <- CRS('+init=epsg:4326') ## Setting default projection
ufo_sp <- spTransform(ufo_locs,
                      CRS("+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs "))
# US region / domain
usa_utm <- spTransform(region, 
                       CRS("+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
```


#### Mesh construction and weights

An example of mesh used for model UFO locations. The location of UFO are used as initial triangulation nodes and use US shape to build a boundary for the points.

```{r mesh, results='asis', fig.width=8, fig.height=8, warning=FALSE, message=FALSE, fig.cap="Mesh constructed using US map."}
ufo_coo <- coordinates(ufo_sp)
mesh <- inla.mesh.2d(loc = ufo_coo, boundary = inla.sp2segment(usa_utm), max.edge = c(75000,200000), cutoff = 50000)
plot(mesh, asp=1)
nv <- mesh$n ## number of vertices in the mesh
n <- nrow(ufo) ## number of observations
```


```{r}
## define the SPDE model
ufo_spde <- inla.spde2.pcmatern(mesh = mesh, 
                                alpha = 2,
                                prior.range = c(0.1, 0.01), # P(range < 0.1) = 0.01
                                prior.sigma = c(0.01, 0.01)) # P(sigma > 0.01) = 0.01
dmesh <- book.mesh.dual(mesh)
usabdy.sp <- SpatialPolygons(list(Polygons(list(Polygon(usa_utm)), ID = "1")))
library(rgeos)
usabdy.sp <- gBuffer(usabdy.sp, byid=TRUE, width=0)
w <- sapply(1:length(dmesh), function(i) {
  if (gIntersects(dmesh[i, ], usabdy.sp)) 
    return(gArea(gIntersection(dmesh[i, ], usabdy.sp)))
  else 
    return(0) 
})
sum(w)
table(w > 0)
```



#### Projector matrix and data stack

```{r}
y.pp <- rep(0:1, c(nv, n))
e.pp <- c(w, rep(0, n))
imat <- Diagonal(nv, rep(1, nv))
lmat <- inla.spde.make.A(mesh, coordinates(ufo_sp))
A.pp <- rbind(imat, lmat)
## the data stack
stk.pp <- inla.stack(
  data = list(y = y.pp, e = e.pp),
  A = list(1, A.pp),
  effects = list(list(b0 = rep(1, nv + n)), list(i = 1:nv)),
  tag = 'pp')
```


#### Fitting the model with no covariate and some results

```{r}
pp.res <- inla(y ~ 0 + b0 + f(i, model = ufo_spde),
               family = 'poisson', 
               data = inla.stack.data(stk.pp),
               control.predictor = list(A = inla.stack.A(stk.pp)),
               control.compute = list(dic = TRUE, cpo = TRUE, waic = TRUE),
               E = inla.stack.data(stk.pp)$e)
## transform to user scale
pp.res.est <- inla.spde.result(inla = pp.res, name = "i", spde = ufo_spde, do.transf = TRUE)

## some results
summary(pp.res)
pp.res$summary.fixed
pp.res$summary.hyperpar
```


```{r res1, results='asis', warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## marginal posterior distribution
par(mfrow=c(2,2), mar=c(3,3,1,0.1), mgp=2:0)
plot(pp.res$marginals.fixed[[1]], type='l', xlab=expression(beta), ylab='Density') 
plot(pp.res.est$marginals.variance.nominal[[1]], type='l', xlab=expression(sigma[x]), ylab='Density')
plot(pp.res.est$marginals.kappa[[1]], type='l', xlab=expression(kappa), ylab='Density')
plot(pp.res.est$marginals.range.nominal[[1]], type='l', xlab='range nominal', ylab='Density')
```

*Posterior distribution for the hyperparameters for $\beta_0$, $\sigma^2$, $\kappa$ and the nominal range.*


#### Projection on a grid

First, we calculate a grid projection from the `mesh` and we project the posterior mean and the posterior standard deviation on the both grid.


```{r rf1, results='asis', warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
library(lattice)
library(gridExtra)
gproj <- inla.mesh.projector(mesh, dims = c(1000, 1000))
g_mean <- inla.mesh.project(gproj, pp.res$summary.random$i$mean)
g_sd <- inla.mesh.project(gproj, pp.res$summary.random$i$sd)
do.call('grid.arrange',
        lapply(list(g_mean, g_sd),
               levelplot, col.regions=terrain.colors(16),
               scales=list(draw=FALSE), xlab='', ylab=''))
```


*The posterior mean and posterior variance for the random field without covariate.*


## Including a covariate on the log-Cox process


#### Adding US population as covariate

* US 2020 Cartographic Boundary Files downloaded from https://www2.census.gov/geo/tiger/GENZ2020/shp/cb_2020_us_county_20m.zip

* 2020 Census Demographic Data By County downloaded from https://www.dataemporium.com/dataset/254/


```{r, warning=FALSE, message=FALSE, }
library(sf)
library(dplyr)

## US 2020 Cartographic Boundary Files
US20 <- st_read("data/cb_2020_us_county_20m/cb_2020_us_county_20m.shp")
## leave out AK, HI, and PR (state FIPS: 02, 15, and 72)
conti_US20 <- US20[!(US20$STATEFP %in% c("02","15","72")), ]

## 2020 Census Demographic Data By County
census20 <- read.csv("data/US_2020_census.csv")
contiguous <- subset(census20, census20$STATE_ABBR != "AK" & census20$STATE_ABBR != "HI" & census20$STATE_ABBR != "PR")
pop20 <- contiguous[, 1:3]

### append total population to US20
conti_US20 <- arrange(conti_US20, NAMELSAD)
pop20 <- arrange(pop20, COUNTY)
conti_US20$TOT_POP <- pop20$TOTAL_POPULATION

### create shape object with state polygons
us_pop20 <- conti_US20
us_pop20 <- as(us_pop20, "Spatial")
us_pop20 <- spTransform(us_pop20, 
                        CRS("+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
pop_mesh <- sp::over(SpatialPoints(mesh$loc[,1:2], proj4string = CRS(proj4string(ufo_sp))), us_pop20)$TOT_POP
pop_obs <- sp::over(ufo_sp, us_pop20)$TOT_POP
covs <- data.frame(pop = c(pop_mesh, pop_obs))
## data stack
stk.cov <- inla.stack(
  data = list(y = y.pp, e = e.pp), 
  A = list(1, A.pp),
  effects = list(list(b0 = rep(1, nv + n), pop = covs$pop), list(i = 1:nv)),
  tag = 'pp')
```


#### Fitting the model and some results


```{r}
pp.cov <- inla(y ~ 0 + b0 + pop + f(i, model = ufo_spde), 
               family = 'poisson', data = inla.stack.data(stk.cov), 
               control.predictor = list(A = inla.stack.A(stk.cov)), 
               control.compute = list(dic = TRUE, cpo = TRUE, waic = TRUE),
               E = inla.stack.data(stk.pp)$e)
## transfer
pp.cov.est <- inla.spde.result(inla = pp.cov, name = "i", spde = ufo_spde, do.transf = TRUE)
## coefficients of the fixed effects
summary(pp.cov)
pp.cov$summary.fixed
pp.cov$summary.hyperpar
```



```{r res2, results='asis', warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
par(mfrow=c(2,2), mar=c(3,3,1,0.1), mgp=2:0) 
plot(pp.cov$marginals.fixed[[1]], type='l', xlab=expression(beta), ylab='Density') 
plot(pp.cov.est$marginals.variance.nominal[[1]], type='l', xlab=expression(sigma[x]), ylab='Density')
plot(pp.cov.est$marginals.kappa[[1]], type='l', xlab=expression(kappa), ylab='Density')
plot(pp.cov.est$marginals.range.nominal[[1]], type='l', xlab='range nominal', ylab='Density')
```

*Posterior distribution for the hyperparameters for $\beta_0$, $\sigma^2$, $\kappa$ and the nominal range.*



#### Projection on a grid


```{r rf2, results='asis', warning=FALSE, message=FALSE, fig.width=10, fig.height=10}
g_mean_cov <- inla.mesh.project(gproj, pp.cov$summary.random$i$mean)
g_sd_cov <- inla.mesh.project(gproj, pp.cov$summary.random$i$sd)
do.call('grid.arrange',
        lapply(list(g_mean_cov, g_sd_cov),
               levelplot, col.regions=terrain.colors(16),
               scales=list(draw=FALSE), xlab='', ylab=''))
```


*The posterior mean and posterior variance for the random field with covariate.*



