---
title: ""
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Overview

Investigating mesh attributes in fitting LGCPs using the INLA-SPDE approach.

```{r triang_properties, eval = TRUE, echo = FALSE, fig.cap="Illustration of the different triangle attributes refered to below"}
points <- data.frame(x = c(1, 2, 3),
                     y = c(1,1, 2.5))
library(deldir)
del <- deldir(points)
source("functions.r")
O <- circum_O(c(1, 1), c(2, 1), c(3, 2.5))
R <- circum_R(c(1, 1), c(2, 1), c(3, 2.5))

iO <- incircle_O(c(1, 1), c(2, 1), c(3, 2.5))
r <- incircle_r(c(1, 1), c(2, 1), c(3, 2.5))
## stuff for point on AC line
b <- iO[2] + (4/3)*iO[1]
x <- (b - 1/4) * 12/25
y <- (3/4)*x + 1/4

## plot
par(xpd = TRUE)
plot(del, wlines = "triang", asp = 1, ylim = c(1,4), xlim = c(0,3),
     pch = 20, lwd = 2)
points(O[1], O[2], pch = 4, lwd = 3)
## circumcircle
plotrix::draw.circle(O[1], O[2], R)
arrows(O[1], O[2], points[1,1], points[1, 2], length = 0, lty = 2)
arrows(O[1], O[2], points[2,1], points[2, 2], length = 0, lty = 2)
arrows(O[1], O[2], points[3,1], points[3, 2], length = 0, lty = 2)
## text
text(points - c(1,1,-1)*0.1,
     labels = c(expression(italic("x")),expression(italic("y")),expression(italic("z"))))
text(O[1],O[2] + 0.15, labels = expression( italic("C")))
text(1.1, 1.75,  labels = expression( italic("R")))
text(1.5, 1.1,  labels = expression("\u2113"[min]))
text(2.6, 2.1, expression(theta))
plotrix::draw.arc(3, 2.5, 0.5, deg1 = 217, deg2 = 236)
## incircle
points(iO[1], iO[2], pch = 20, col = "grey")
plotrix::draw.circle(iO[1], iO[2], r, border = "grey")
arrows(iO[1], iO[2], points[1,1], points[1, 2], length = 0, lty = 2, col = "grey")
arrows(iO[1], iO[2], points[2,1], points[2, 2], length = 0, lty = 2, col = "grey")
arrows(iO[1], iO[2], points[3,1], points[3, 2], length = 0, lty = 2, col = "grey")
arrows(iO[1], iO[2], x, y, length = 0, col = "grey")
text(1.7, 1.35,  labels = expression( italic("r")), col = "grey")
text(iO[1],iO[2] - 0.1, labels = expression( italic("c")), col = "grey")
```


**Note**: The radius-edge ratio of a polygon $\tau$ is given by $re = \frac{R}{l_{\text{min}}}$, where R is the circumradius and $l_{\text{min}}$ is the length of the shortest edge. The radius ratio $rr = \frac{r}{R}$, where $r$ is the inradius.

We can visualise each triangle measure:

```{r hores__re, eval = TRUE, echo = FALSE}
url <- "https://gist.githubusercontent.com/cmjt/9a5e43cce69a3babb129b0a448e65752/raw/8653d11cb6ace667a3f5683982ae82366a3ac5d2/horse.csv"
horse <- read.csv(url)
source("functions.r")
library(sp)
library(INLA)
library(ggplot2)
library(patchwork)
sp <- sp::SpatialPolygons(list(sp::Polygons(list(sp::Polygon(horse)), '0')))
bound <- inla.sp2segment(sp)

m1 <- INLA::inla.mesh.2d(boundary = bound,
                         max.edge = 2)
m3 <- INLA::inla.mesh.create(horse, bound = bound)
m1_tris <- get_triag_attributes(m1)
m3_tris <- get_triag_attributes(m3)

## mesh attribute plots

rr3 <- ggplot(m3_tris$sf, aes(fill = m3_tris$triangles$rr)) +
    geom_sf(colour = NA) + theme_void() +
    scale_fill_continuous(name = "Radius Ratio", limits = c(0, 0.5))


rr1 <- ggplot(m1_tris$sf, aes(fill = m1_tris$triangles$rr)) +
    geom_sf(colour = NA) + theme_void() +
    scale_fill_continuous(name = "Radius Ratio", limits = c(0, 0.5))

(rr3 / rr1) + plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect")

```



## Files

 + `functions.r` contains a number of function to calculate different metrics measures and attributes. The main function
 `get_triag_attributes()` takes a `inla.mesh.2d()` object and returns a named list of length three: 
     
     + `edges` a dataframe of mesh segments  
     + `triangles` a dataframe with, for each triangle in the triangulation, the circumcircle and incircle radii,
     the associated centroid locations, and the radius-edge ratio & radius ratio,
     + `angles` is a 3 x nvert dataframe of all triangle interior angles
     
 + `data/ufo.csv` subset of data sourced from the [UFO Sightings Map](https://www.arcgis.com/apps/webappviewer/index.html?id=ddda71d5211f47e782b12f3f8d06246e) collated [here](https://data.world/timothyrenner/ufo-sightings#). Data for 2020 only and excludes sightings in Alaska and Hawaiʻi.
 
```{r ufo}
ufo <- readr::read_csv("data/ufo.csv")
ufo
```
